Real-Time Variational Density Forecasts

### 2017-03-03 Confirmation/confirmation_V2.tex
updates
more presentation changes

Mean Field Variational Bayes (MFVB) has origins in physics (\citealp{Amari1985}; \citealp{Chandler1987}) and has since developed a large literature in machine learning, notably developed in \citet{Jordan1999} and \citet{Ghahramani2000}. MFVB restricts the class of distributions from which to select $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ to the set of factorisable distributions,
\begin{equation}
\label{mf1}
q(\boldsymbol{\theta} | \boldsymbol{\lambda}) = \prod_{i=1}^k q_i(\theta_i | \boldsymbol{\lambda_i}).
\end{equation}

@@

\citet{Ranganath2014} showed that a Monte Carlo estimate of the derivative of the ELBO can be given by
\begin{equation}
\label{SVB2}
\nabla_{\lambda}\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{S}\sum_{s=1}^{S} \nabla_{\boldsymbol{\lambda}} [\ln(q(\boldsymbol{\theta}_s | \boldsymbol{\lambda}^{(m)}))] (\ln (p(y_{1:T}, \boldsymbol{\theta}_s)) - \ln(q(\boldsymbol{\theta}_s | \boldsymbol{\lambda}^{(m)})))
\end{equation}
where $s = 1, \dots, S$ indicates simulations from $q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)})$.

\citet{Kingma2014} introduces the re-parameterisation trick, where an auxillary noise variable $\boldsymbol{\epsilon}$ and differentiable transformation $f(\cdot, \cdot)$ are introduced such that $\boldsymbol{\theta} = f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}$. Examples include a location scale transformation from a standard normal $\epsilon$ or an inverse-CDF transformation from a uniform$(0, 1) \epsilon$. In this case (\ref{SVB2}) can be replaced with a lower variance estimator,
\begin{equation}
\label{reparameterise}
\nabla_{\lambda}\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{S}\sum_{s=1}^{S} \nabla_{\boldsymbol{\lambda}} \left( \ln (p(y_{1:T}, f(\boldsymbol{\epsilon}_s, \boldsymbol{\lambda}^{(m)})) - \ln(q(f(\boldsymbol{\epsilon}_s, \boldsymbol{\lambda}^{(m)})) \right)
\end{equation}
where $s = 1, \dots, S$ now indicates simulations from $q(\boldsymbol{\epsilon})$. This transformation removes the restriction that $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ belong to a distribution where the score function can be evaluated.

>>>>

Mean Field Variational Bayes (MFVB) has origins in physics \citep{Peterson1987} and has since developed a large literature in machine learning, notably developed in \citet{Jordan1999} and \citet{Ghahramani2000}. MFVB restricts the class of distributions from which to select $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ to the set of factorisable distributions,
\begin{equation}
\label{mf1}
q(\boldsymbol{\theta} | \boldsymbol{\lambda}) = \prod_{i=1}^k q_i(\theta_i | \boldsymbol{\lambda_i}).
\end{equation}

@@

\citet{Ranganath2014} showed that a Monte Carlo estimate of the derivative of the ELBO can be given by
\begin{equation}
\label{SVB2}
\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{S}\sum_{s=1}^{S} \nabla_{\boldsymbol{\lambda}} [\ln(q(\boldsymbol{\theta}_s | \boldsymbol{\lambda}^{(m)}))] (\ln (p(y_{1:T}, \boldsymbol{\theta}_s)) - \ln(q(\boldsymbol{\theta}_s | \boldsymbol{\lambda}^{(m)})))
\end{equation}
where $s = 1, \dots, S$ indicates simulations from $q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)})$.

\citet{Kingma2014} introduces the re-parameterisation trick, where an auxillary noise variable $\boldsymbol{\epsilon}$ and differentiable transformation $f(\cdot, \cdot)$ are introduced such that $\boldsymbol{\theta} = f(\boldsymbol{\epsilon}, \boldsymbol{\lambda}$. Examples include a location scale transformation from a standard normal $\epsilon$ or an inverse-CDF transformation from a uniform$(0, 1) \epsilon$. In this case (\ref{SVB2}) can be replaced with a lower variance estimator,
\begin{equation}
\label{reparameterise}
\nabla_{\boldsymbol{\lambda}}\mathcal{L}(q(\boldsymbol{\theta} | \boldsymbol{\lambda}^{(m)}), y_{1:T}) \approx \frac{1}{S}\sum_{s=1}^{S} \nabla_{\boldsymbol{\lambda}} \left( \ln (p(y_{1:T}, f(\boldsymbol{\epsilon}_s, \boldsymbol{\lambda}^{(m)})) - \ln(q(f(\boldsymbol{\epsilon}_s, \boldsymbol{\lambda}^{(m)})) \right)
\end{equation}
where $s = 1, \dots, S$ now indicates simulations from $q(\boldsymbol{\epsilon})$. This transformation removes the restriction that $q(\boldsymbol{\theta} | \boldsymbol{\lambda})$ belong to a distribution where the score function can be evaluated.

### 2017-03-03 Confirmation/presentations.tex
more presentation changes

\documentclass{beamer}

@@

\title[Real-Time Variational Density Forecasts]{Real-Time Variational Density Forecasts}
\author[Nathaniel Tomasetti]{Nathaniel Tomasetti}
\date{ }
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\begin{frame}
\section{Motivation}
\frametitle{Motivation}
\begin{itemize}
\item There is an increasing demand for forecasts to be of the entire predictive density of the variable of interest.
\item Many time series of interest can operate on a very short time-frame.
\item We aim to provide density forecasts for these high frequency time-series by updating posterior and predictive density estimates as each data point is observed.
\item The algorithm use should scale to multivariate observations driven by complex models with large datasets.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Gradient Ascent}
\begin{itemize}
\item Make steps of the form
\begin{equation}
\lambda^{(m+1)} = \lambda^{(m)} + \rho^{(m)} \nabla_{\lambda} f(\lambda^{(m)})
\end{equation}
until $| f(\lambda^{(m+1)}) - f(\lambda^{(m)}) | < \epsilon$.
\end{itemize}
\begin{figure}[h]
\centering
<<gradascent, out.width='0.4\\textwidth', error=FALSE, message=FALSE, warning=FALSE, echo=FALSE>>=
library(ggplot2)
x = rep(0, 20)
x[1] = -2.5
for(i in 1:19){
  deriv = -x[i]*dnorm(x[i])
  x[i+1] = x[i] + 23 * deriv / i
}
y2 = dnorm(x)
library(ggplot2)
xsup = seq(-3, 1.5, 0.01)
y = dnorm(xsup, 0, 1)

ggplot() + geom_line(aes(xsup, y)) + geom_path(aes(x, y2), colour = "red") + labs(y = "Gaussian Density", x = expression(lambda)) + theme_bw()
@
\caption{Red steps show the path a variable $\lambda^{(m)}$ took while finding the maximum of a standard normal distribution.}
\label{fig:GradAscent}
\end{figure}
\end{frame}

@@

\begin{frame}
\section{Empirical Application}
\frametitle{Electricity Load Forecasting}
Motivation - Electricity 
\end{frame}

>>>>

\documentclass{beamer}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}

@@

\title[Real-Time Variational Density Forecasts]{Real-Time Variational Density Forecasts}
\author[Nathaniel Tomasetti]{Nathaniel Tomasetti}
\date{ }
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\begin{frame}
\section{Motivation}
\frametitle{Motivation}
\begin{itemize}
\item There is an increasing demand for forecasts of the entire predictive density of the variable of interest.
\item Many time series of interest can operate on a very short time-frame.
\item We aim to provide density forecasts for these high frequency time-series by updating posterior and predictive density estimates as each data point is observed.
\item The algorithm used should scale to multivariate observations driven by complex models with large datasets.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Gradient Ascent}
\begin{itemize}
\item Make steps of the form
\begin{equation}
\lambda^{(m+1)} = \lambda^{(m)} + \rho^{(m)} \nabla_{\lambda} f(\lambda^{(m)})
\end{equation}
until $| f(\lambda^{(m+1)}) - f(\lambda^{(m)}) | < \epsilon$, with $\rho^{(m)}, m = 1, 2, \dots$ that satisfies the \citet{Robbins1951} conditions.
\end{itemize}
\begin{figure}[h]
\centering
\includegraphics[scale = 0.4]{gradientascent}
\caption{Red steps show the path a variable $\lambda^{(m)}$ took while finding the maximum of a standard normal distribution.}
\label{fig:GradAscent}
\end{figure}
\end{frame}

@@

\begin{frame}
\section{Empirical Application}
\frametitle{Electricity Load Forecasting}
\begin{itemize}
\item Victorian electricity prices are extremely volatile, and are updated every five minutes.
\item Prices spike electricity load crosses a particular threshold.
\item If we can provide density forecasts for load, we can predict how likely it is for prices to spike in the next period.
\begin{figure}[h]
\centering
\includegraphics[scale = 0.3]{RRP_timeseries.png}
\caption{Half hourly price data for Victoria from January 1 2011 to December 31 2015.}
\label{fig:rrpplot}
\end{figure}
\end{itemize}
\end{frame}
